---
title: "Join_LSOA_IMD_Infras"
author: "Eugeni Vidal"
date: "03/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(sf)
library(tidyverse)
library(tmap)
```

```{r, eval=FALSE}
# Load data
## Load infras data
infras = readRDS("Data/ways_all.Rds")
## Create length_m variable
infras$length_m = as.numeric(st_length(infras))
```

```{r setup, include=FALSE}
# Get IMD data
knitr::opts_chunk$set(message = TRUE)
u <- "https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/467774/File_7_ID_2015_All_ranks__deciles_and_scores_for_the_Indices_of_Deprivation__and_population_denominators.csv"
d <- readr::read_csv(u)
names(d) <- snakecase::to_snake_case(names(d))
names(d)[1] <- "geo_code"

# Pct package download
z <- pct::get_pct(layer = "z", geography = "lsoa", national = TRUE)

# Join datasets
zm <- dplyr::inner_join(z, d)

## Load zm data
zm <- zm %>% 
  st_transform(27700) # add geometry
# plot(las$geometry) # ckeck visually
# zm = zm %>% filter(str_detect(lad17cd, "E")) # select only England
```

```{r}
# Drop variables not interesting
names(zm)
zm <- zm[,c(1:6,128:129,135,186)]
colnames(zm)[7] <- "imd_decile"
```

# Aggregate lsoa per IMD and add slope and perc_cyc

```{r}
# Aggregate per IMD
lads_imd <- zm %>% 
  group_by(lad11cd, imd_decile) %>% 
  summarise(av_slope = mean(avslope_perc_u10km), perc_cyc = sum(bicycle)/sum(all))
```

### Infrastructure per lad

```{r}
### Infrastructure per lsoa
# Extract categories of infrastructure datasets based on https://wiki.openstreetmap.org/wiki/Bicycle#cite_note-anyroad-1. 
## Extract cycleways highway dataset
i_cycle_tracks = filter(infras, highway == "cycleway") 
## Extract cycleways left or right dataset
i_cycle_lanes = filter(infras, (cycleway.left != "no" | cycleway.right != "no") & highway != "cycleway")
```

```{r}
# Aggregate infrastructure variables to zm
# lads_imd$length_infras = aggregate(infras["length_m"], lads_imd, FUN = sum)$length_m
# Aggregate cycleways highway
lads_imd$length_cycle_tracks = aggregate(i_cycle_tracks["length_m"], lads_imd, FUN = sum)$length_m
## Replace NA by 0
lads_imd$length_cycle_tracks[is.na(lads_imd$length_cycle_tracks)] = 0
# Aggregate cycleways left or right
lads_imd$length_cycle_lanes = aggregate(i_cycle_lanes["length_m"], lads_imd, FUN = sum)$length_m
## Replace NA by 0
lads_imd$length_cycle_lanes[is.na(lads_imd$length_cycle_lanes)] = 0
```

```{r message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, paged.print=FALSE}
# Create infrastructure indicators
lads_imd$area <- st_area(lads_imd)
## Convert area m^2 to km^2
lads_imd$km2 <- lads_imd$area/1000000
## Cycling segregated infrastructure per area
lads_imd$cycle_tracks_km2 = (lads_imd$length_cycle_tracks/1000)/(lads_imd$km2)
## Cycling non segregated infrastructure per area
lads_imd$cycle_lanes_km2 = (lads_imd$length_cycle_lanes/1000)/(lads_imd$km2)
## Total cycle-friendly infrastructure per area.
lads_imd$total_cycle_inf_km2 = ((lads_imd$length_cycle_tracks + lads_imd$length_cycle_lanes)/1000)/(lads_imd$km2)
```

### Cyclability

```{r, include = FALSE} 
# Create the bikeability weights from dodgr package
weighting_profiles = dodgr::weighting_profiles$weighting_profiles
bike_profile = weighting_profiles %>%
  filter(name == "bicycle") %>%
  select(highway = way, value)
#bike_profile
## Join this file with the bikeability weights
infras = left_join(infras, bike_profile)
```

```{r}
# Shouldn't I calculate this indicator by lenght?
# Aggregate cyclability indicator
lads_imd$cyclability = aggregate(infras["value"], lads_imd, na.rm = TRUE, FUN = mean)$value
```

### Volume of traffic

```{r}
# Create a weighted volume column by cycling flow
## Shold I use the function weighted.mean() or multiply by dutch scenario as Robin suggested?
infras$volume_w = infras$aadt * infras$pct.census
# Aggregate volume_w indicator
lads_imd$volume_w = aggregate(infras["volume_w"], lads_imd, na.rm = TRUE, FUN = mean)$volume_w
```

### Quietness

```{r}
# Sholdn't I calculate quietness by lenght?

# Aggregate cyclability indicator
lads_imd$quietness = aggregate(infras["quietness"], lads_imd, na.rm = TRUE, FUN = mean)$quietness
```

```{r}
# save test data
saveRDS(lads_imd, "Data/dirty_env_lsoa.Rds")
```

```{r}
lads_imd <- readRDS("Data/dirty_env_lsoa.Rds")
```

```{r}
names(lads_imd)
lads_imd <- lads_imd[,c(1:4,128:129,135,186:196)]
```

```{r}
# save test data
saveRDS(lads_imd, "Data/env_lsoa.Rds")
```

```{r}
# Aggregate per IMD
lads_imd <- lads_imd %>% 
  group_by(lad11cd, index_of_multiple_deprivation_imd_decile_where_1_is_most_deprived_10_of_lso_as) %>% 
  summarise(length_cycle_tracks = sum(length_cycle_tracks), length_cycle_lanes = sum(length_cycle_lanes), av_slope = mean(avslope_perc_u10km))
```


### Data distribution

- Data visualisation. The total_cycle_friendly_inf_km2 data is very skewed to the left. However, the cyclability index looks normalised. 

```{r}
# Infrastructure data very skezed to the left. 
hist(lads_imd$total_cycle_inf_km2)
#hist(log(lads_imd$total_cycle_friendly_inf_km2))
```

```{r}
hist(lads_imd$cyclability)
```

```{r}
hist(lads_imd$volume_w)
```

```{r}
hist(lads_imd$quietness)
```

The skeweness of the first indicator is because cycling infrastructure is very low and tend to be concentrated in urban areas. But it might be also because OSM collaborations are higher in urban than in rural areas. How can we check this? One solution could be to analyse only specific urban areas - those in where the level of contributions is enough to make the data reliable.

The cyclability index data is normalised because it takes into account all types of infrastructure, not only cycle-friendly infrastructure. It looks much more reliable.
